Rust for Scientific Computing

### 2025-08-20 main.tex
add more ideas
define preliminary structure of paper

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}

@@

\title{Rust for High-Performance Computing}

@@

@@

%###################################################################################################
\section{Ergonomics }
\label{section:debugging}

@@

%---------------------------------------------------------------------------------------------------
\subsection{Numerics}
\label{subsection:numerics}

\begin{enumerate}
    \item ODE solvers~\cite{Renevey2024}
    \item PDE solvers
    \item Fluid Dynamics: Salva (2D \& 3D)~\cite{Crozet2024}
    \item Physics Engine: Rapier (2D \& 3D)~\cite{Crozet2025}
\end{enumerate}

%---------------------------------------------------------------------------------------------------
\subsection{Storage}
\label{subsection:storage}

\begin{enumerate}
    \item serde ecosystem~\cite{TolnaySerde2024,TolnaySerdeJson2025}
\end{enumerate}

%---------------------------------------------------------------------------------------------------
\subsection{Debugging}
\label{subsection:debugging}

%---------------------------------------------------------------------------------------------------
\subsection{GPGPU Programming}
\label{subsection:gpgpu}

\begin{enumerate}
    \item wgpu~\cite{Fitzgerald2025}
\end{enumerate}

%###################################################################################################
\section{Examples}
\label{section:examples}

\begin{enumerate}
    \item Physics~\cite{}
    \item Biology~\cite{Pleyer2024}
    \item Bioinformatics~\cite{}
\end{enumerate}

\bibliographystyle{IEEEtran}
\bibliography{references}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}
\usepackage[backend=biber,style=ieee]{biblatex}

\addbibresource{references.bib}

@@

\title{Rust for Scientific Computing}

@@

\pagebreak
\tableofcontents
\pagebreak

@@

@@

\begin{enumerate}
    \item Linear Algebra \& Arrays (nalgebra, ndarray, ...)
    \item Statistics \& Data Analysis (statrs, ndarray-stats, smartcore, ...)
    \item Symbolic \& other Math (num, rug, speciyl, symcc)
    \item ODE solvers~\cite{Renevey2024}, Diffsol
    \item PDE solvers
    \item Fluid Dynamics: Salva (2D \& 3D)~\cite{Crozet2024}
    \item Physics Engine: Rapier (2D \& 3D)~\cite{Crozet2025}
    \item Meshing (honeycomb)
    \item Plotting (plotters, plotlib, gnuplot bindings, vtk-rs)
    \item Utilities (rayon, lapack)
    \item GPU (rust-cuda, wgpu)
    \item Datastructures
    \item Optimization
\end{enumerate}

\subsection{Foundational Components}
\subsubsection{Linear Algebra \& Arrays}
(nalgebra, ndarray) (dense, sparse matrices, solvers)

\subsubsection{Numerical Methods}
(floating point arithmetic, etc.)

\subsubsection{Statistics \& Probability}
(statsrs, ndarray-stats, smartcore)

\subsubsection{Symbolic \& Algebraic Computing}
(num, rug, speciyl, symcc, enzyme)

\subsubsection{Datastructures}
(graphs, etc.)

\subsection{Computational Methods}
\subsubsection{Optimization}
\subsubsection{Differential Equations}
(ODEs, PDEs)

\subsubsection{Physics Engines}
(Fluid Dynamics, Manybody)

\subsubsection{Numerical Integration \& Differentiation}
(quadrature, monte-carlo, symbolic diff, enzyme)

\subsubsection{(Probabilistic Algorithms)}

\subsection{High-Performance \& Scalable Computing}
\subsubsection{Parallel Computing}
(rayon)

\subsubsection{GPU \& Accelerators}
(rust-cuda, wgpu~\cite{Fitzgerald2025})

\subsubsection{Sparse \& Structured Computations}
(large databases)

\subsubsection{Debugging}
(Rust Errors as values, tracing, logging) ecosystem

\subsubsection{Storage}
(serde, json, xml, ron, ...)

\subsection{Visualization}
\subsubsection{Plotting (2D)}
\subsubsection{3D Rendering}
\subsubsection{Interactive}
(dioxus)

%###################################################################################################
\section{Applications}
\label{section:applications}

\subsection{Physics \& Engineering}
\subsection{Biology \& Chemistry}
\cite{Pleyer2024}, Computational Oncology~\cite{Köster2025}, Proteomics~\cite{Anechitoaie2024}
\subsection{Climate \& Earth Sciences}
\subsection{Economics \& Finance}
\subsection{Social Sciences}

%###################################################################################################
\section{Discussion}
\label{section:discussion}

%###################################################################################################
\section{Conclusion}
\label{section:conclusion}

\printbibliography
